### 2013-09-22 image_proc_methods.tex
final draft methods

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
Axial T2W MR slices were manually segmented using the polygon tool on ITK SNAP
using separate labels for the peripheral zone (PZ), central gland (CG) and
anterior fibromuscular stroma (AFS). The gland was segmented from base to apex.
The base was identified below the bladder and subsequent images were segmented
until the last slice with visible prostatic tissue was identified caudally. The
CG, PZ and AFS were segmented independently according to their well-established
anatomical characteristics on
T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogenous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous and lower signal intensity as well as its
location. Although not readily visible on every case, the AFS was identified by
its low T2 signal intensity and its location anterior to the central gland.

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base to apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, PZ and AFS were segmented
independently according to their well-established anatomical characteristics on
T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogenous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous and lower signal intensity as well as its
location. Although not readily visible on every case, the AFS was identified by
its low T2 signal intensity and its location anterior to the central gland.

ARFI images were segmented using a similar procedure to the MR images.

\emph{CO-AUTHOR QUESTION: CAN WE EASILY DELINEATE OUR SEGMENTATION CRITERIA?}

Segmented image stacks were imported into 3D Slicer (v4.3.0) and 3D
models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

\begin{table}[h!]
\centering
\caption{3D model volume rendering parameters}
\begin{tabular}{ll}
{\bf Parameter} & {\bf Value} \\ \hline
Decimation & 0.1 \\
Smoothing Algorithm & Laplacian \\
Smoothing  & 70.0 \\
Joint Smoothing & Enabled \\
\end{tabular}
\end{table}

The 3D slicer models were used to render volume estimates of the PZ and CG,
with the sum of PZ and CG representing the total prostate gland volume.  AFS
volume estimates in select MR cases were included in the total prostate gland
volume estimates.  Orthogonal tri-axial measurements in the lateral-to-lateral,
axex-to-base, and anterior-to-posterior dimensions were made in 3D slicer.

### 2013-09-22 mr_methods.tex
final draft methods

\subsection{Study Inclusion Criteria}
Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
this study.  Inclusion criteria were undergoing complete pelvic MRI with
endorectal coil for detection of prostate cancer, including multiplanar
T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
prostatectomy.  Patients with previous treatments of PCa or benign prostatic
hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
study.  All patients enrolled in this study provided written informed consent.

\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}

\subsection{MR Imaging}
All MR imaging was performed on one of two 3.0 Tesla MR scanners (General
Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare,
Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
Indianola, PA), as well as multi-channel surface coils.  Imaging sequences
included thin-section (3 mm section thickness) fast spin echo T2-weighted
images in the coronal, axial and sagittal planes.  Diffusion weighted images
were obtained using multiple b-values and calculation of ADC maps was also
performed.  Dynamic contrast enhanced MR sequences were obtained after
administration of a weight-based dose of extracellular MR contrast agent with
4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
table with full MR parameters) Prostates were radically removed using a da
Vinci Surgical System (Intuitive Surgical\textregistered, Inc., Sunnyvale, CA).
After exision, the prostates were weighed and measured, formalin fixed for at
least 24 hours without being cut, and then processed for whole mount histology.

>>>>

\subsection{MR Imaging}
All MR imaging was performed on one of two 3.0 Tesla MR scanners (General
Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare,
Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
Indianola, PA), as well as multi-channel surface coils.  Imaging sequences
included thin-section (3 mm section thickness) fast spin echo T2-weighted
images in the coronal, axial and sagittal planes.  Diffusion weighted images
were obtained using multiple b-values and calculation of ADC maps was also
performed.  Dynamic contrast enhanced MR sequences were obtained after
administration of a weight-based dose of extracellular MR contrast agent with
4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
table with full MR parameters) Prostates were radically removed using a da
Vinci Surgical System (Intuitive Surgical\textregistered, Inc., Sunnyvale, CA).
